---
title: "damageDEMOleafdata"
author: "Shelley Bennett"
date: "3/18/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Load packages
```{r}
library(nlme)
library(emmeans)
library(performance)
library(tidyverse)
library(codyn)
library(lme4)
library(pscl)
library(effects)
library(glmmTMB)
library(car) #Anova test for mixed models, for this code of the genpois model form glmmTMB
library(DHARMa)

```


Random stuff that I need
```{r}
#run this for sum of squares
options(contrasts=c('contr.sum','contr.poly')) #run this first, important to make sure your sum of squares work right

##function for making graphs with summary statistics
barGraphStats <- function(data, variable, byFactorNames) {
  count <- length(byFactorNames)
  N <- aggregate(data[[variable]], data[byFactorNames], FUN=length)
  names(N)[1:count] <- byFactorNames
  names(N) <- sub("^x$", "N", names(N))
  mean <- aggregate(data[[variable]], data[byFactorNames], FUN=mean)
  names(mean)[1:count] <- byFactorNames
  names(mean) <- sub("^x$", "mean", names(mean))
  sd <- aggregate(data[[variable]], data[byFactorNames], FUN=sd)
  names(sd)[1:count] <- byFactorNames
  names(sd) <- sub("^x$", "sd", names(sd))
  preSummaryStats <- merge(N, mean, by=byFactorNames)
  finalSummaryStats <- merge(preSummaryStats, sd, by=byFactorNames)
  finalSummaryStats$se <- finalSummaryStats$sd / sqrt(finalSummaryStats$N)
  return(finalSummaryStats)
}  
```

load data
```{r}
damage <- read.csv("~/Dropbox (Smithsonian)/SERC_damageDemo/damage_2021_combined.csv")
plots <- read.csv("~/Dropbox (Smithsonian)/SERC_damageDemo/damagedemoplots.csv")
treenum <- read.csv("~/Dropbox (Smithsonian)/SERC_damageDemo/treenum.csv")
```
Amy load data
```{r}

setwd("C:/Users/hrusk/Dropbox (Smithsonian)/SERC_damageDEMO")
damage <- read.csv("damage_2021_combined2.csv")
plots <- read.csv("damagedemoplots2.csv")

```

merge data
```{r}
#merge dataframes
damage2 <- damage %>%
  merge(plots, by = "Plot")
  #merge(treenum, by = "Tree_num")

#filter for only sapling and mature
damage1 <- damage2 %>%
  filter(Age_class == "Sapling" | Age_class == "Mature")

damage1$Age_class <- factor(damage1$Age_class, levels = c("Sapling", "Mature"))
damage2$Age_class <- factor(damage2$Age_class, levels = c("Seedling", "Sapling", "Mature"))

damageMature <- damage1 %>%
  filter(Age_class == "Mature")

damageSapling <- damage1 %>%
  filter(Age_class == "Sapling")
```

stuff that is wrong or doesn't work
```{r}
#tried running an ANOVA without repeated measures
summary(insectherbModel <- aov(Percent_total_herbivory~as.factor(Species)*as.factor(Age_class)*as.factor(plot_age_class)*as.factor(Sample_period), data=damage1))
anova(insectherbModel)


#this is the model we want but doesn't currently run
summary(insectherbModel <- lme(Percent_total_herbivory~as.factor(Species)*as.factor(Age_class)*as.factor(plot_age_class), data=damage1, random=~1|uniqueplant_num, correlation=corCompSymm(form=~1|uniqueplant_num), control=lmeControl(returnObject = T)))
##this model doesn't work because we included plot age class in the fixed effects and there aren't all interactions in all plot age classes (i.e not all species at each plant age class in all plot age classes)

##trying again with nesting
summary(insectherbModel <- lme(Percent_total_herbivory~as.factor(Species)*as.factor(Age_class)*as.factor(plot_age_class), data=damage1, random=~1|plot_age_class/Plot/Age_class/Tree_num, correlation=corCompSymm(form=~1|plot_age_class/Plot/Tree_num), control=lmeControl(returnObject = T)))
```


######Total Percent Herbivory Model######
```{r}
damage1$Species <- as.factor(damage1$Species)
damage1$Age_class <- as.factor(damage1$Age_class)
damage1$Sample_period <- as.factor(damage1$Sample_period)

summary(insectherbModel <- lme(Percent_total_herbivory~Species*Age_class*Sample_period, data=damage1, random=~1|plot_age_class/Plot/Tree_num, correlation=corCompSymm(form=~1|plot_age_class/Plot/Tree_num), control=lmeControl(returnObject = T)))

anova.lme(insectherbModel, type = 'sequential')
emmeans(insectherbModel, pairwise~as.factor(Species)*as.factor(Age_class), adjust="tukey")
```

looking at above model
```{r}
plot(allEffects(insectherbModel))
```

checking data
```{r}
hist(damage1$Percent_total_herbivory)
qqnorm(damage1$Percent_total_herbivory)
```


overdispersion function
```{r}
overdisp_fun <- function(model) {
     rdf <- df.residual(model)
     rp <- residuals(model,type="pearson")
     Pearson.chisq <- sum(rp^2)
     prat <- Pearson.chisq/rdf
     pval <- pchisq(Pearson.chisq, df=rdf, lower.tail=FALSE)
     c(chisq=Pearson.chisq,ratio=prat,rdf=rdf,p=pval)}

overdisp_fun(insectherbModel)

```


model to account for overdispersion/non-nomral distribution
doesn't work for Shelley, need "OpenMP"
```{r}
damage1$Percent_total_herbivory <- as.integer(as.numeric(damage1$Percent_total_herbivory))

insectherbModeloverdisp <-glmmTMB(Percent_total_herbivory~ Species*Age_class*Sample_period+(1|plot_age_class/Plot/Tree_num), data = damage1, family = genpois)

##model checking##
model_check <- simulateResiduals(insectherbModeloverdisp)
plot(model_check)

#Amy also ran...#
testZeroInflation(insectherbModeloverdisp) #not significant, good
testUniformity(insectherbModeloverdisp) #not significant, good
testOutliers(insectherbModeloverdisp) #not significant, good

drop1(insectherbModeloverdisp, test = 'Chisq') #preferred way to test model, but doesn't look at main effects

#Model output:
#Percent_total_herbivory ~ Species * Age_class * Sample_period + 
#    (1 | plot_age_class/Plot/Tree_num)
#                                Df   AIC    LRT Pr(>Chi)   
#<none>                             19926                   
#Species:Age_class:Sample_period  2 19933 10.522 0.005191 **
#---
#Signif. codes:  0 ‘***’ 0.001 ‘**’ 0.01 ‘*’ 0.05 ‘.’ 0.1 ‘ ’ 1

Anova(insectherbModeloverdisp) #Wald's test not prefferred for model type, but gives main effects

#Analysis of Deviance Table (Type II Wald chisquare tests)

#Response: Percent_total_herbivory
#                                   Chisq Df Pr(>Chisq)    
#Species                         102.6882  2  < 2.2e-16 ***
#Age_class                         0.3208  1    0.57111    
#Sample_period                    28.4578  1  9.576e-08 ***
#Species:Age_class                 5.6101  2    0.06050 .  
#Species:Sample_period            63.7586  2  1.429e-14 ***
#Age_class:Sample_period           1.9274  1    0.16504    
#Species:Age_class:Sample_period   9.0769  2    0.01069 *  
#---
#Signif. codes:  0 ‘***’ 0.001 ‘**’ 0.01 ‘*’ 0.05 ‘.’ 0.1 ‘ ’ 1


emmeans(insectherbModeloverdisp, pairwise~as.factor(Species)*as.factor(Age_class), adjust="tukey")

```



non-normal model with seedling data and without sample period
```{r}

damage2$Percent_total_herbivory <- as.integer(as.numeric(damage2$Percent_total_herbivory))

insectherbModeloverdisp2 <-glmmTMB(Percent_total_herbivory~ Species*Age_class+(1|plot_age_class/Plot/Tree_num), data = damage2, family = genpois)

##model checking##
model_check2 <- simulateResiduals(insectherbModeloverdisp2)
plot(model_check2) #within-group deviations from uniformity significant, 
#Levene Test for homogeneity of variance signficant

testZeroInflation(insectherbModeloverdisp2) #not significant, good
testUniformity(insectherbModeloverdisp2) #not significant, good
testOutliers(insectherbModeloverdisp2) #not significant, good

drop1(insectherbModeloverdisp2, test = 'Chisq')
#Model:
#Percent_total_herbivory ~ Species * Age_class + (1 | plot_age_class/Plot/Tree_num)
#                  Df   AIC    LRT  Pr(>Chi)    
#<none>               20854                     
#Species:Age_class  4 20866 20.342 0.0004274 ***
#---
#Signif. codes:  0 ‘***’ 0.001 ‘**’ 0.01 ‘*’ 0.05 ‘.’ 0.1 ‘ ’ 1

Anova(insectherbModeloverdisp2)

#Analysis of Deviance Table (Type II Wald chisquare tests)

#Response: Percent_total_herbivory
#                    Chisq Df Pr(>Chisq)    
#Species            87.067  2  < 2.2e-16 ***
#Age_class         150.575  2  < 2.2e-16 ***
#Species:Age_class  16.278  4   0.002668 ** 
#---
#Signif. codes:  0 ‘***’ 0.001 ‘**’ 0.01 ‘*’ 0.05 ‘.’ 0.1 ‘ ’ 1

emmeans(insectherbModeloverdisp2, pairwise~as.factor(Species)*as.factor(Age_class), adjust="tukey")
```

microbial models for saplings and mature
```{r}
damage1$Percent_microbial <- as.integer(as.numeric(damage1$Percent_microbial))

insectherbModeloverdisp3 <-glmmTMB(Percent_microbial~ Species*Age_class*Sample_period+(1|plot_age_class/Plot/Tree_num), data = damage1, family = genpois)

##model checking##
model_check3 <- simulateResiduals(insectherbModeloverdisp3)
plot(model_check3) #KS test = deviation significant, might not be of concern? this is better than poisson distribution

#Amy also ran...#
testZeroInflation(insectherbModeloverdisp3) #not significant, good
testUniformity(insectherbModeloverdisp3) #significant deviation
testOutliers(insectherbModeloverdisp3) #not significant, good

drop1(insectherbModeloverdisp3, test = 'Chisq')
#Model:
#Percent_microbial ~ Species * Age_class * Sample_period + (1 | 
#    plot_age_class/Plot/Tree_num)
#                                Df   AIC    LRT  Pr(>Chi)    
#<none>                             26498                     
#Species:Age_class:Sample_period  2 26526 31.942 1.158e-07 ***
#---
#Signif. codes:  0 ‘***’ 0.001 ‘**’ 0.01 ‘*’ 0.05 ‘.’ 0.1 ‘ ’ 1

Anova(insectherbModeloverdisp3)
#Analysis of Deviance Table (Type II Wald chisquare tests)

#Response: Percent_microbial
#                                    Chisq Df Pr(>Chisq)    
#Species                          216.6381  2  < 2.2e-16 ***
#Age_class                          1.0601  1    0.30320    
#Sample_period                   1765.5411  1  < 2.2e-16 ***
#Species:Age_class                  0.9832  2    0.61164    
#Species:Sample_period            192.9121  2  < 2.2e-16 ***
#Age_class:Sample_period            5.8862  1    0.01526 *  
#Species:Age_class:Sample_period   31.6474  2  1.342e-07 ***
#---
#Signif. codes:  0 ‘***’ 0.001 ‘**’ 0.01 ‘*’ 0.05 ‘.’ 0.1 ‘ ’ 1

emmeans(insectherbModeloverdisp3, pairwise~as.factor(Species)*as.factor(Age_class), adjust="tukey")

```

microbial models for seedlings without sampling period
```{r}
damage2$Percent_microbial <- as.integer(as.numeric(damage2$Percent_microbial))

insectherbModeloverdisp4 <-glmmTMB(Percent_microbial~ Species*Age_class+(1|plot_age_class/Plot/Tree_num), data = damage2, family = genpois)

##model checking##
model_check4 <- simulateResiduals(insectherbModeloverdisp4)
plot(model_check4) #KS test: significant deviation
#within-group deviations from uniformity significant, 
#Levene Test for homogeneity of variance significant

testZeroInflation(insectherbModeloverdisp4) #not significant, good
testUniformity(insectherbModeloverdisp4) #significant deviation
testOutliers(insectherbModeloverdisp4) #not signfificant, good

drop1(insectherbModeloverdisp4, test = 'Chisq')
#Model:
#Percent_microbial ~ Species * Age_class + (1 | plot_age_class/Plot/Tree_num)
#                  Df   AIC    LRT  Pr(>Chi)    
#<none>               30659                     
#Species:Age_class  4 30732 81.004 < 2.2e-16 ***
#---
#Signif. codes:  0 ‘***’ 0.001 ‘**’ 0.01 ‘*’ 0.05 ‘.’ 0.1 ‘ ’ 1

Anova(insectherbModeloverdisp4)

#Analysis of Deviance Table (Type II Wald chisquare tests)

#Response: Percent_microbial
#                    Chisq Df Pr(>Chisq)    
#Species           191.951  2  < 2.2e-16 ***
#Age_class         194.177  2  < 2.2e-16 ***
#Species:Age_class  88.239  4  < 2.2e-16 ***
#---
#Signif. codes:  0 ‘***’ 0.001 ‘**’ 0.01 ‘*’ 0.05 ‘.’ 0.1 ‘ ’ 1

emmeans(insectherbModeloverdisp4, pairwise~as.factor(Species)*as.factor(Age_class), adjust="tukey")
```



graphs
```{r}


theme_set(theme_bw())
theme_update(axis.title.x=element_text(size=24, vjust=-0.35), axis.text.x=element_text(size=20, color = "black"),
             axis.title.y=element_text(size=24, angle=90, vjust=0.7), axis.text.y=element_text(size=20, color= "black"),
             plot.title = element_text(size=24, vjust=2),
             panel.grid.major=element_blank(), panel.grid.minor=element_blank(),
             legend.title=element_blank(), legend.text=element_text(size=18), panel.border=element_rect(color="black", fill = NA, size = 1))

ggplot(data=barGraphStats(data=damage1, variable="Percent_total_herbivory", byFactorNames=c("Species", "Age_class")), aes(x=Species, y=mean, fill=Age_class)) +
  geom_bar(stat='identity', position=position_dodge()) +
  geom_errorbar(aes(ymin=mean-se, ymax=mean+se), width=0.2, position=position_dodge(0.9))+
  theme(legend.title=element_blank())+
  theme_bw()+
  scale_fill_manual(values=c("darkolivegreen3", "darkolivegreen4"))+
  ylab('Mean Percent Herbivore Damage\n')+
  xlab(element_blank())+
  expand_limits(y=4) +
  annotate("text", x= 0.77, y = 3.5, label= "a", size = 6)+
  annotate("text", x= 1.23, y = 3.7, label= "a", size = 6)+
  annotate("text", x= 1.77, y = 2.7, label= "b", size = 6)+
  annotate("text", x= 2.23, y = 2.2, label= "b", size = 6)+
  annotate("text", x= 2.77, y = 3.4, label= "ab", size = 6)+
  annotate("text", x= 3.23, y = 3.8, label= "a", size = 6)+
  theme(legend.title=element_blank(),legend.text=element_text(size=16), axis.text.x=element_text(size=16, color = "black"), axis.title.y=element_text(size=18), axis.text.y=element_text(size=16, color= "black"), panel.grid.major=element_blank(), panel.grid.minor=element_blank())+
  expand_limits(y=9)
```

###herbivory with seedlings###
```{r}
summary(insectherbModelseed <- lme(Percent_total_herbivory~Species*Age_class, data=damage2, random=~1|plot_age_class/Plot/Age_class/Tree_num, correlation=corCompSymm(form=~1|plot_age_class/Plot/Age_class/Tree_num), control=lmeControl(returnObject = T)))

anova.lme(insectherbModelseed, type = 'sequential')
emmeans(insectherbModelseed, pairwise~as.factor(Species)*as.factor(Age_class), adjust="tukey")
```


graphing with seedlings
```{r}
ggplot(data=barGraphStats(data=damage2, variable="Percent_total_herbivory", byFactorNames=c("Species", "Age_class")), aes(x=Species, y=mean, fill=Age_class)) +
  geom_bar(stat='identity', position=position_dodge()) +
  geom_errorbar(aes(ymin=mean-se, ymax=mean+se), width=0.2, position=position_dodge(0.9))+
  theme(legend.title=element_blank())+
  theme_bw()+
  scale_fill_manual(values=c("darkolivegreen3", "darkolivegreen4", "darkolivegreen"))+
  ylab('Mean Percent Herbivore Damage\n')+
  xlab(element_blank())+
  expand_limits(y=4) +
  theme(legend.title=element_blank(),legend.text=element_text(size=16), axis.text.x=element_text(size=16, color = "black"), axis.title.y=element_text(size=18), axis.text.y=element_text(size=16, color= "black"), panel.grid.major=element_blank(), panel.grid.minor=element_blank())+
  expand_limits(y=9)
```

#####Microbial model#########

```{r}
summary(microbialModel <-lme(Percent_microbial~as.factor(Species)*as.factor(Age_class)*as.factor(Sample_period), data=damage1, random=~1|plot_age_class/Plot, correlation=corCompSymm(form=~1|plot_age_class/Plot/plant_num), control=lmeControl(returnObject = T)))

anova.lme(microbialModel, type = 'sequential')
emmeans(microbialModel, pairwise~as.factor(Species)*as.factor(Age_class), adjust="tukey")
```

```{r}

ggplot(data=barGraphStats(data=damage1, variable="Percent_microbial", byFactorNames=c("Species", "Age_class")), aes(x=Species, y=mean, fill=Age_class)) +
  geom_bar(stat='identity', position=position_dodge()) +
  geom_errorbar(aes(ymin=mean-se, ymax=mean+se), width=0.2, position=position_dodge(0.9))+
  theme_bw()+
  scale_fill_manual(values=c("darkolivegreen3", "darkolivegreen4"))+
  ylab('Mean Percent Microbial Damage\n')+
  xlab(element_blank())+
  annotate("text", x= 0.77, y = 2.7, label= "c", size = 6)+
  annotate("text", x= 1.23, y = 2.8, label= "c", size = 6)+
  annotate("text", x= 1.77, y = 7.9, label= "b", size = 6)+
  annotate("text", x= 2.23, y = 9.3, label= "a", size = 6)+
  annotate("text", x= 2.77, y = 9.5, label= "ab", size = 6)+
  annotate("text", x= 3.23, y = 8.7, label= "b", size = 6)+
  theme(legend.title=element_blank(),legend.text=element_text(size=16), axis.text.x=element_text(size=16, color = "black"), axis.title.y=element_text(size=18), axis.text.y=element_text(size=16, color= "black"), panel.grid.major=element_blank(), panel.grid.minor=element_blank())+
  expand_limits(y=9)

```

graph with seedlings
```{r}
ggplot(data=barGraphStats(data=damage2, variable="Percent_microbial", byFactorNames=c("Species", "Age_class")), aes(x=Species, y=mean, fill=Age_class)) +
  geom_bar(stat='identity', position=position_dodge()) +
  geom_errorbar(aes(ymin=mean-se, ymax=mean+se), width=0.2, position=position_dodge(0.9))+
  theme_bw()+
  scale_fill_manual(values=c("darkolivegreen3", "darkolivegreen4", "darkolivegreen"))+
  ylab('Mean Percent Microbial Damage\n')+
  xlab(element_blank())+
  theme(legend.title=element_blank(),legend.text=element_text(size=16), axis.text.x=element_text(size=16, color = "black"), axis.title.y=element_text(size=18), axis.text.y=element_text(size=16, color= "black"), panel.grid.major=element_blank(), panel.grid.minor=element_blank())+
  expand_limits(y=9)
```

######Regressions###


practice regression graphs
```{r}
herbivory_lm <- lm(Percent_total_herbivory ~ plot_age*Species*Age_class*Sample_period, data = damage1)
summary(herbivory_lm)
#Age_class, sampling period

predictedherbivory <- predict(herbivory_lm, interval = "confidence", level = 0.95) 

full_herbivory <- data.frame(damage1, predictedherbivory)


ggplot(full_herbivory, aes(x = plot_age, y = Percent_total_herbivory))+
  geom_point(aes(color= Species))+
  geom_line(aes(x = plot_age, y = fit)) +
  geom_ribbon(aes(x = plot_age, ymin = lwr, ymax = upr), alpha = 0.3) +
  #annotate("text", x = 14, y = 290, label = "y = -7.35x + 161.33")+
  #annotate("text", x = 14, y = 270, label = "italic(R) ^ 2 == 0.14", parse = TRUE)+
  ylab('Percent herbivory')+
  xlab('Plot age')
 
ggplot(subset(damage1, Sample_period == 2 & Age_class == "Mature" & plot_age < 200), aes(x=plot_age, y=log10(Percent_total_herbivory))) +
  geom_point(aes(color=Species)) + 
  scale_color_manual(values = c("#D39200",  "#93AA00", "#00B9E3")) +
  geom_smooth(aes(color=Species, fill=Species), size=1, method=lm) +
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
        panel.background = element_blank(), axis.line = element_line(colour = "black"))
  



number_ticks <- function(n) {function(limits) pretty(limits, n)}
sp.labs <- c("Jul", "Oct")
names(sp.labs) <- c("1", "2")

# Create the plot
#p + facet_grid(
  #dose ~ supp, 
 # labeller = labeller(dose = dose.labs, supp = supp.labs))

ggplot(damage1, aes(x=plot_age, y=log10(Percent_total_herbivory))) +
  geom_point(aes(color=Species)) + 
  scale_color_manual(values = c("yellow3",  "chartreuse4", "darkorchid3")) +
  geom_smooth(aes(color=Species, fill=Species), size=1, method=lm) +
  scale_x_continuous(breaks=number_ticks(4)) +
  facet_grid(rows = vars(Age_class), cols = vars(Sample_period), labeller = labeller(Sample_period = sp.labs))+
  theme(strip.text.x = element_text(size = 14, color = "black"), strip.text.y = element_text(size = 14, color = "black"))+
  ylab("log10(Percent Herbivore Damage)")+
  xlab("Plot Age")+
  theme(legend.title=element_blank(),legend.text=element_text(size=16), axis.text.x=element_text(size=16, color = "black"), axis.title.y=element_text(size=18), axis.text.y=element_text(size=16, color= "black"), axis.title.x=element_text(size=18), panel.grid.major=element_blank(), panel.grid.minor=element_blank())

ggplot(damage1, aes(x=plot_age, y=log10(Percent_microbial))) +
  geom_point(aes(color=Species)) + 
  scale_color_manual(values = c("yellow3",  "chartreuse4", "darkorchid3")) +
  geom_smooth(aes(color=Species, fill=Species), size=1, method=lm) +
  scale_x_continuous(breaks=number_ticks(4)) +
  facet_grid(rows = vars(Age_class), cols = vars(Sample_period), labeller = labeller(Sample_period = sp.labs))+
  theme(strip.text.x = element_text(size = 14, color = "black"), strip.text.y = element_text(size = 14, color = "black"))+
  ylab("log10(Percent Microbial Damage)")+
  xlab("Plot Age")+
  theme(legend.title=element_blank(),legend.text=element_text(size=16), axis.text.x=element_text(size=16, color = "black"), axis.title.y=element_text(size=18), axis.text.y=element_text(size=16, color= "black"), axis.title.x=element_text(size=18), panel.grid.major=element_blank(), panel.grid.minor=element_blank())



#mature, sample period 1
ggplot(subset(damage1, Sample_period == 1 & Age_class == "Mature"), aes(x=plot_age, y=log10(Percent_total_herbivory))) +
  geom_point(aes(color=Species)) + 
  scale_color_manual(values = c("yellow3",  "chartreuse4", "darkorchid3")) +
  geom_smooth(aes(color=Species, fill=Species), size=0.5, method=lm) +
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
        panel.background = element_blank(), axis.line = element_line(colour = "black"))

#mature sample period 2
ggplot(subset(damage1, Sample_period == 2 & Age_class == "Mature"), aes(x=plot_age, y=log10(Percent_total_herbivory))) +
  geom_point(aes(color=Species)) + 
  scale_color_manual(values = c("yellow3",  "chartreuse4", "darkorchid3")) +
  geom_smooth(aes(color=Species, fill=Species), size=0.5, method=lm) +
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
        panel.background = element_blank(), axis.line = element_line(colour = "black"))

#sapling sample period 1
ggplot(subset(damage1, Sample_period == 1 & Age_class == "Sapling"), aes(x=plot_age, y=log10(Percent_total_herbivory))) +
  geom_point(aes(color=Species)) + 
  scale_color_manual(values = c("yellow3",  "chartreuse4", "darkorchid3")) +
  geom_smooth(aes(color=Species, fill=Species), size=1, method=lm) +
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
        panel.background = element_blank(), axis.line = element_line(colour = "black"))

#sapling sample period 2
ggplot(subset(damage1, Sample_period == 2 & Age_class == "Sapling"), aes(x=plot_age, y=log10(Percent_total_herbivory))) +
  geom_point(aes(color=Species)) + 
  scale_color_manual(values = c("yellow3",  "chartreuse4", "darkorchid3")) +
  geom_smooth(aes(color=Species, fill=Species), size=1, method=lm) +
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
        panel.background = element_blank(), axis.line = element_line(colour = "black"))
```


####facet regression graphs####

```{r}
number_ticks <- function(n) {function(limits) pretty(limits, n)}
sp.labs <- c("Jul", "Oct")
names(sp.labs) <- c("1", "2")

# Create the plot
#p + facet_grid(
  #dose ~ supp, 
 # labeller = labeller(dose = dose.labs, supp = supp.labs))

ggplot(damage1, aes(x=plot_age, y=log10(Percent_total_herbivory))) +
  geom_point(aes(color=Species)) + 
  scale_color_manual(values = c("yellow3",  "chartreuse4", "darkorchid3")) +
  geom_smooth(aes(color=Species, fill=Species), size=1, method=lm) +
  scale_x_continuous(breaks=number_ticks(4)) +
  facet_grid(rows = vars(Age_class), cols = vars(Sample_period), labeller = labeller(Sample_period = sp.labs))+
  theme(strip.text.x = element_text(size = 14, color = "black"), strip.text.y = element_text(size = 14, color = "black"))+
  ylab("log10(Percent Herbivore Damage)")+
  xlab("Plot Age")+
  theme(legend.title=element_blank(),legend.text=element_text(size=16), axis.text.x=element_text(size=16, color = "black"), axis.title.y=element_text(size=18), axis.text.y=element_text(size=16, color= "black"), axis.title.x=element_text(size=18), panel.grid.major=element_blank(), panel.grid.minor=element_blank())

ggplot(damage1, aes(x=plot_age, y=log10(Percent_microbial))) +
  geom_point(aes(color=Species)) + 
  scale_color_manual(values = c("yellow3",  "chartreuse4", "darkorchid3")) +
  geom_smooth(aes(color=Species, fill=Species), size=1, method=lm) +
  scale_x_continuous(breaks=number_ticks(4)) +
  facet_grid(rows = vars(Age_class), cols = vars(Sample_period), labeller = labeller(Sample_period = sp.labs))+
  theme(strip.text.x = element_text(size = 14, color = "black"), strip.text.y = element_text(size = 14, color = "black"))+
  ylab("log10(Percent Microbial Damage)")+
  xlab("Plot Age")+
  theme(legend.title=element_blank(),legend.text=element_text(size=16), axis.text.x=element_text(size=16, color = "black"), axis.title.y=element_text(size=18), axis.text.y=element_text(size=16, color= "black"), axis.title.x=element_text(size=18), panel.grid.major=element_blank(), panel.grid.minor=element_blank())





ggplot(damage2, aes(x=plot_age, y=log10(Percent_total_herbivory))) +
  geom_point(aes(color=Species)) + 
  scale_color_manual(values = c("yellow3",  "chartreuse4", "darkorchid3")) +
  geom_smooth(aes(color=Species, fill=Species), size=1, method=lm) +
  scale_x_continuous(breaks=number_ticks(4)) +
  facet_grid(rows = vars(Age_class), cols = vars(Sample_period), labeller = labeller(Sample_period = sp.labs))+
  theme(strip.text.x = element_text(size = 14, color = "black"), strip.text.y = element_text(size = 14, color = "black"))+
  ylab("log10(Percent Herbivore Damage)")+
  xlab("Plot Age")+
  theme(legend.title=element_blank(),legend.text=element_text(size=16), axis.text.x=element_text(size=16, color = "black"), axis.title.y=element_text(size=18), axis.text.y=element_text(size=16, color= "black"), axis.title.x=element_text(size=18), panel.grid.major=element_blank(), panel.grid.minor=element_blank())

ggplot(damage2, aes(x=plot_age, y=log10(Percent_microbial))) +
  geom_point(aes(color=Species)) + 
  scale_color_manual(values = c("yellow3",  "chartreuse4", "darkorchid3")) +
  geom_smooth(aes(color=Species, fill=Species), size=1, method=lm) +
  scale_x_continuous(breaks=number_ticks(4)) +
  facet_grid(rows = vars(Age_class), cols = vars(Sample_period), labeller = labeller(Sample_period = sp.labs))+
  theme(strip.text.x = element_text(size = 14, color = "black"), strip.text.y = element_text(size = 14, color = "black"))+
  ylab("log10(Percent Microbial Damage)")+
  xlab("Plot Age")+
  theme(legend.title=element_blank(),legend.text=element_text(size=16), axis.text.x=element_text(size=16, color = "black"), axis.title.y=element_text(size=18), axis.text.y=element_text(size=16, color= "black"), axis.title.x=element_text(size=18), panel.grid.major=element_blank(), panel.grid.minor=element_blank())

```


####stacked bar charts#####

```{r}
herbivorywide <- damage1 %>%
  select (Plot, Species, Age_class, Leaf_num, Sample_period, Percent_chewed, Percent_mined, Percent_skeletonized, Percent_gall, plot_age, plot_age_class)

herbivorylong <- gather(herbivorywide, Herbivory_type, Percent_damage, Percent_chewed:Percent_gall, factor_key=TRUE)
herbivorylong

herbivorylong$Sample_period <- as.factor(herbivorylong$Sample_period)

ggplot(data=barGraphStats(data=herbivorylong, variable="Percent_damage", byFactorNames=c("Species", "Sample_period", "Herbivory_type")), aes(x=Sample_period, y=mean, fill=Herbivory_type))+
  geom_bar(position="stack", stat="identity")+
  facet_wrap(~Species)+
  ylab('Mean Percent Damage')+
  xlab('Sample Period')+
  theme(legend.title=element_blank())
  #facet_grid(rows = vars(Age_class), cols = vars(Species))

herbivorylongMature <- herbivorylong %>%
  filter(Age_class == "Mature")

ggplot(data=barGraphStats(data=herbivorylongMature, variable="Percent_damage", byFactorNames=c("Species", "Sample_period", "Herbivory_type")), aes(x=Sample_period, y=mean, fill=Herbivory_type))+
  geom_bar(position="stack", stat="identity")+
  scale_fill_discrete(labels = c("Chewing", "Mining", "Skeletonizing", "Galls"))+
  facet_wrap(~Species)+
  theme(strip.text.x = element_text(size = 14, color = "black"))+
  ylab('Mean Percent Herivore Damage on Mature Trees\n')+
  xlab('Sample Period')+
  theme(legend.title=element_blank())+
  theme(legend.title=element_blank(),legend.text=element_text(size=16), axis.text.x=element_text(size=16, color = "black"), axis.title.y=element_text(size=18), axis.text.y=element_text(size=16, color= "black"), axis.title.x=element_text(size=18), panel.grid.major=element_blank(), panel.grid.minor=element_blank())+
  scale_x_discrete(labels = c("Jul", "Oct"))

herbivorylongSapling <- herbivorylong %>%
  filter(Age_class == "Sapling")

ggplot(data=barGraphStats(data=herbivorylongSapling, variable="Percent_damage", byFactorNames=c("Species", "Sample_period", "Herbivory_type")), aes(x=Sample_period, y=mean, fill=Herbivory_type))+
  geom_bar(position="stack", stat="identity", width = 1)+
  scale_fill_discrete(labels = c("Chewing", "Mining", "Skeletonizing", "Galls"))+
  facet_wrap(~Species)+
  theme(strip.text.x = element_text(size = 14, color = "black"))+
  ylab('Mean Percent Herbivore Damage on Saplings\n')+
  xlab('Sample Period')+
  theme(legend.title=element_blank())+
  theme(legend.title=element_blank())+
  theme(legend.title=element_blank(),legend.text=element_text(size=16), axis.text.x=element_text(size=16, color = "black"), axis.title.y=element_text(size=18), axis.text.y=element_text(size=16, color= "black"), axis.title.x=element_text(size=18), panel.grid.major=element_blank(), panel.grid.minor=element_blank())+
  scale_x_discrete(labels = c("Jul", "Oct"))

damage1$Sample_period <- as.factor(damage1$Sample_period)

ggplot(data=barGraphStats(data=damage1, variable = "Percent_microbial", byFactorNames=c("Species", "Sample_period")), aes(x=Sample_period, y=mean))+
  geom_bar(position="dodge", stat="identity")+
  facet_wrap(~Species)+
  ylab('Mean Percent Microbial Damage')+
  xlab('Sample Period')


ggplot(data=barGraphStats(data=damageMature, variable = "Percent_microbial", byFactorNames=c("Species", "Sample_period")), aes(x=Sample_period, y=mean))+
  geom_bar(position="dodge", stat="identity",  fill="chocolate2")+
  facet_wrap(~Species)+
  theme(strip.text.x = element_text(size = 14, color = "black"))+
  ylab('Mean Percent Microbial Damage on Mature Trees\n')+
  xlab('Sample Period')+
  theme(legend.title=element_blank(),legend.text=element_text(size=16), axis.text.x=element_text(size=16, color = "black"), axis.title.y=element_text(size=18), axis.text.y=element_text(size=16, color= "black"), axis.title.x=element_text(size=18), panel.grid.major=element_blank(), panel.grid.minor=element_blank())+
  scale_x_discrete(labels = c("Jul", "Oct"))+
  expand_limits(y=15)

ggplot(data=barGraphStats(data=damageSapling, variable = "Percent_microbial", byFactorNames=c("Species", "Sample_period")), aes(x=Sample_period, y=mean))+
  geom_bar(position="dodge", stat="identity", fill="chocolate2")+
  facet_wrap(~Species)+
  theme(strip.text.x = element_text(size = 14, color = "black"))+
  ylab('Mean Percent Microbial Damage\n')+
  xlab('Sample Period')+
  theme(legend.title=element_blank())+
  theme(legend.title=element_blank(),legend.text=element_text(size=16), axis.text.x=element_text(size=16, color = "black"), axis.title.y=element_text(size=18), axis.text.y=element_text(size=16, color= "black"), axis.title.x=element_text(size=18), panel.grid.major=element_blank(), panel.grid.minor=element_blank())+
  scale_x_discrete(labels = c("Jul", "Oct"))+
  expand_limits(y=15)

```


stacked graphs for seedlings
```{r}

herbivorywideSeedlings <- damage2 %>%
  select (Plot, Species, Age_class, Leaf_num, Sample_period, Percent_chewed, Percent_mined, Percent_skeletonized, Percent_gall, plot_age, plot_age_class)

herbivorylongSeedlings <- gather(herbivorywideSeedlings, Herbivory_type, Percent_damage, Percent_chewed:Percent_gall, factor_key=TRUE)
herbivorylong

herbivorylongSeedlings$Sample_period <- as.factor(herbivorylongSeedlings$Sample_period)

herbivorylongSeedlingsSeed <- herbivorylongSeedlings%>%
  filter(Age_class == "Seedling")

ggplot(data=barGraphStats(data=herbivorylongSeedlingsSeed, variable="Percent_damage", byFactorNames=c("Species", "Sample_period", "Herbivory_type")), aes(x=Sample_period, y=mean, fill=Herbivory_type))+
  geom_bar(position="stack", stat="identity")+
  scale_fill_discrete(labels = c("Chewing", "Mining", "Skeletonizing", "Galls"))+
  facet_wrap(~Species)+
  theme(strip.text.x = element_text(size = 14, color = "black"))+
  ylab('Mean Percent Herivore Damage\n')+
  xlab('Sample Period')+
  theme(legend.title=element_blank())+
  theme(legend.title=element_blank(),legend.text=element_text(size=16), axis.text.x=element_text(size=16, color = "black"), axis.title.y=element_text(size=18), axis.text.y=element_text(size=16, color= "black"), axis.title.x=element_text(size=18), panel.grid.major=element_blank(), panel.grid.minor=element_blank())+
  scale_x_discrete(labels = c("Jul", "Oct"))
```

######original code from Kim####
```{r}
##code from Kim below
#below is the model
summary(insectherbModel <- lme(avg_perc_herbivory~as.factor(species)*as.factor(ind_age)*as.factor(stand_age), #typical model with percent herbivory as a dependent variable and species, individual age, and stand age as explanatory variables
                               data=herbivoryData, #the dataframe the model is being run on
                               random=~1|plot, #random statement for which plot the plant came from
                               correlation=corCompSymm(form=~date|plot/plant), #compound symmetry correlation structure (good for herbivory, which isn't necessarily increasing through time), saying that plants are nested within plots, which were sampled multiple dates
                               control=lmeControl(returnObject=T))) #I forget what this does, but I think it is important to keep

#this checks that the model ran ok (checks assumptions, might not work exactly right on this model though)
check_model(insectherbModel)

#this gives you your p-values
anova.lme(insectherbModel, type='sequential') 


#this gives you the mean and standard errors for each treatment
emmeans(insectherbModel, pairwise~as.factor(stand_age), adjust="tukey")




#same model with a sqrt transformation to make the data normal
summary(insectherbModel <- lme(sqrt(avg_perc_herbivory)~as.factor(species)*as.factor(ind_age)*as.factor(stand_age), #note how avg_perc_herbivory is square root transformed in this one
                               data=herbivoryData, 
                               random=~1|plot, 
                               correlation=corCompSymm(form=~date|plot/plant),
                               control=lmeControl(returnObject=T)))
check_model(insectherbModel)
anova.lme(insectherbModel, type='sequential') 
back.emmeans(emmeans(insectherbModel, pairwise~as.factor(stand_age), adjust="tukey"), transform='sqrt') #back.emmeans will backtransform the values for you 
```

